Cross-validation of regression models: mixed model examples (edited v3)

### 2025-03-04 scratch/JSS-submission/Resubmission/response-to-reviewer.Rmd

---
title: "Point-by-Point Response to the Reviewer"
author: "John Fox and Georges Monette"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

2.  "What is the purpose of using insight::get_data on page 5? The insight package is also mentioned on page 31. I believe it would be helpful to include a brief explanation of why it is being used."

As suggested, we added a brief explanation of the advantage of using the **insight** package, which provides a robust way to access model components for many classes of statistical models.

3.  "Does the package allow for the addition of an uncertainty interval to the plots in Figure 2? I think that including a measure of uncertainty would be useful."

Yes, it does. As we explained on p. 5 of the original paper in discussing the `confint` argument to `cv()`, "`confint`, whether or not to compute a confidence interval for the CV criterion, defaulting to `TRUE` if there are at least 400 cases." The `Auto` data used in the preliminary example has fewer than 400 cases. When there are confidence limits in the `"cv"` objects, the `plot()` method used for Fig. 2 draws confidence intervals. We've added a footnote (fn. 5) explaining this behaviour and have also in describing the `confint` argument pointed to the appendix of the paper for an explanation of the $n \ge 400$ rule for computing confidence intervals by default.

@@

4.  "In Section 4.1, six pages of the text are devoted to transformations before the actual CV content is addressed. Would it be possible/appropriate to shorten the discussion on transformations?"

We've compressed this material by removing the code and graphs for the scatterplot matrix and C+R plots for the transformed data, shortening the paper by 2 pages. We'd prefer not to abbreviate the section further: The purpose of the section is to show how to automate a complex model-selection method, so we think that it's important to explain how the method is applied "manually" before automating it.

@@

5.  "On page 21, when you state “We proceed to transform the numeric predictors in the Auto regression towards multivariate normality”, I assume that the transformations being applied are marginal rather than joint, as joint transformations would cause the covariates to lose their interpretation. If so, then perhaps it would be worth clarifying that powerTransform adjusts the margins to achieve multivariate Gaussianity."

Yes, the predictors are transformed individually with the object of making their joint distribution as close to multinormal as possible. We have clarified this point in the resubmission.

6.  "On page 29, when computing `mse(Auto\$mpg, exp(fitted(m.step)))`, I think that you could use the relationship between the normal and log-normal distributions to correct for bias."

Yes, one could compute the MSE on the log-scale for both models (which we think is what the reviewer is getting at here). This is a good point, and has a more general implication: It's necessary to compare models on the same response scale, but that need not be the original response scale. We've added to what was footnote 12 on p. 25 of the original submission (now footnote 13 on p. 23) to explain this option (crediting the reviewer).

### 2025-03-04 scratch/JSS-submission/Resubmission/Fox-Monette-cv-package.Rmd
complete draft JSS resubmission and response to reviewer

cv.auto.loo <- cv(mlist, data = Auto, k = "loo")
cv.auto.loo[2]
```
The \code{models()} function takes an arbitrary number of regression models as its arguments, which are optionally named, to create a \code{"modList"} object. Alternatively, as here, we can supply a pre-constructed, optionally named, list of models as the single argument to \code{models()}. To visualize the results, comparing the models, we invoke the \code{plot()} method for the \code{"cvModList"} objects returned by \code{cv()} (see Figure \ref{fig:polynomial-regression-CV-graph-2}):

@@

@@

The \code{powerTransform()} function in the \pkg{car} package transforms variables towards multivariate normality by a generalization of Box and Cox's maximum-likelihood-like approach [@BoxCox:1964]. Several "families" of power transformations can be used, including the original Box-Cox family (which is the default), simple powers (and roots), and two adaptations of the Box-Cox family to data that may include negative values and zeros: the Box-Cox-with-negatives family and the Yeo-Johnson family; see @Weisberg:2014 [Chap. 8] and @FoxWeisberg:2019 [Chap. 3] for details. We proceed to transform the numeric predictors in the \code{Auto} regression towards multivariate normality: 
```{r Auto-transform}
num.predictors <- c("displacement", "horsepower", "weight", 
                    "acceleration")
tr.x <- powerTransform(Auto[, num.predictors])
summary(tr.x)
```
We then apply the (rounded) transformations---all, as it turns out, logs (i.e., "zeroth" powers)---to the data and re-estimate the model:
```{r Auto-with-transformed-predictors}
A <- Auto
powers <- tr.x$roundlam
for (pred in num.predictors){
  A[, pred] <- bcPower(A[, pred], lambda = powers[pred])
}
m <- update(m.auto, data = A)
```

Having transformed the predictors towards multivariate normality, we now consider whether there's evidence for transforming the response (using \code{powerTransform()} for Box and Cox's original method), also obtaining a log transformation:
```{r Auto-Box-Cox}
summary(powerTransform(m))
m <- update(m, log(mpg) ~ .)
```

The transformed numeric variables are much better-behaved (cf., Figure \ref{fig:Auto-explore-scatterplot-matrix}, on page \pageref{fig:Auto-explore-scatterplot-matrix}, and Figure \ref{fig:Auto-transformed-scatterplot-matrix}):
```{r Auto-transformed-scatterplot-matrix}
#| out.width = "95%",
#| fig.align = 'center',
#| fig.height = 5.5,
#| fig.cap = "Scatterplot matrix for the transformed numeric variables in the \\code{Auto} data."
scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
                  + acceleration, 
                  smooth=list(spread = FALSE), data = A, pch = ".")
```

And the partial relationships in the model fit to the transformed data are much more nearly linear (cf., Figure \ref{fig:Auto-explore-cr-plots}, on page \pageref{fig:Auto-explore-cr-plots}, and Figure \ref{fig:Auto-transformed-cr-plots}):

```{r Auto-transformed-cr-plots}
#| out.width = "70%",
#| fig.align = 'center',
#| fig.height = 4.25,
#| fig.cap = "Component + residual plots for the regression model fit to the transformed \\code{Auto} data."
crPlots(m, terms = ~ displacement + horsepower + weight + acceleration, 
        pch = ".", ylab = "C+R", las = 2)
```

@@

[^medAbsErr]: That's slightly uncomfortable given the skewed distribution of \code{mpg}. An alternative is to use a robust measure of model lack-of-fit, such as the median absolute error instead of the mean-squared error, employing the \code{medAbsErr()} function from the \pkg{cv} package. The median absolute error, however, cannot be expressed as a casewise average (see Appendix A.2).

@@

[^times-10]: As before, we ask \code{microbenchmark()} to report relative timings, and also as before, we set \code{times = 10} rather than the default \code{times = 100}, with the consequence that we fail to detect a statistically reliable difference between using the Woodbury matrix identity and using hatvalues for the \code{cv()} computations.

>>>>

cv.auto.loo <- cv(mlist, data = Auto, k = "loo")
cv.auto.loo[2]
```
The \code{models()} function takes an arbitrary number of regression models as its arguments, which are optionally named, to create a \code{"modList"} object. Alternatively, as here, we can supply a pre-constructed, optionally named, list of models as the single argument to \code{models()}. To visualize the results, comparing the models, we invoke the \code{plot()} method for the \code{"cvModList"} objects returned by \code{cv()} (see Figure \ref{fig:polynomial-regression-CV-graph-2}):[^plot-confint]

@@

[^plot-confint]: The objects \code{cv.auto.10} and \code{cv.auto.loo} do not include confidence intervals for the CV MSE estimates because, as we explained, there are fewer than 400 cases in the \code{Auto} data. Consequently, confidence intervals are not shown in the graphs in Figure \ref{fig:polynomial-regression-CV-graph-2}. We can force `cv()` to compute confidence intervals by adding the argument `confint = TRUE`; we invite the reader to do this and to verify that the resulting confidence intervals are very broad, as was the case for the quadratic model fit to the \code{Auto} data in the preceding section.

@@

The \code{powerTransform()} function in the \pkg{car} package transforms variables towards multivariate normality by a generalization of Box and Cox's maximum-likelihood-like approach [@BoxCox:1964]. Although the multinormal likelihood is maximized, the variables are transformed individually. Several "families" of power transformations can be used, including the original Box-Cox family (which is the default), simple powers (and roots), and two adaptations of the Box-Cox family to data that may include negative values and zeros: the Box-Cox-with-negatives family and the Yeo-Johnson family; see @Weisberg:2014 [Chap. 8] and @FoxWeisberg:2019 [Chap. 3] for details. We proceed to transform the numeric predictors in the \code{Auto} regression towards multivariate normality: 
```{r Auto-transform}
num.predictors <- c("displacement", "horsepower", "weight", 
                    "acceleration")
tr.x <- powerTransform(Auto[, num.predictors])
summary(tr.x)
```
We then apply the (rounded) transformations---all, as it turns out, logs (i.e., "zeroth" powers)---to the data and re-estimate the model:
```{r Auto-with-transformed-predictors}
A <- Auto
powers <- tr.x$roundlam
for (pred in num.predictors){
  A[, pred] <- bcPower(A[, pred], lambda = powers[pred])
}
m <- update(m.auto, data = A)
```

Having transformed the predictors towards multivariate normality, we now consider whether there's evidence for transforming the response (using \code{powerTransform()} for Box and Cox's original method), also obtaining a log transformation:
```{r Auto-Box-Cox}
summary(powerTransform(m))
m <- update(m, log(mpg) ~ .)
```
The transformed numeric variables are much better-behaved and the partial relationships in the model fit to the transformed data are much more nearly linear (as we invite the reader to verify by redrawing the scatterplot matrix for the transformed variables and the C+R plots for the regression fit to the transformed data).

@@

[^medAbsErr]: That's slightly uncomfortable given the skewed distribution of \code{mpg}. An alternative is to use a robust measure of model lack-of-fit, such as the median absolute error instead of the mean-squared error, employing the \code{medAbsErr()} function from the \pkg{cv} package. The median absolute error, however, cannot be expressed as a casewise average (see Appendix A.2). Yet another possibility, suggested to us by a reviewer of an earlier version of this paper, is to compute the MSE for both models on the log-response scale rather than on the original scale of the response. This suggestion is quite general, in that the essential point is to use the \emph{same} response scale for all models to be compared, not necessarilty the \emph{original} scale of the response.

@@

[^times-10]: As before, we ask \code{microbenchmark()} to report relative timings, and also as before, we set \code{times = 10} rather than the default \code{times = 100}.
